### 2022-07-10 Journal Version/journal-version.tex
Added files with the new text

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}

\usepackage{epsfig}
\usepackage{subcaption}
\usepackage{calc}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{multicol}
\usepackage{pslatex}
\usepackage{apalike}

\usepackage[scaled=0.8]{helvet}    % Less huge \textsf{functionName}
\usepackage[misc,geometry]{ifsym} % for letter symbol
\usepackage{graphicx}
\usepackage[dvipsnames]{xcolor}
\usepackage{tikz}
\usetikzlibrary{trees,snakes,arrows}
\usetikzlibrary{shapes,chains}
\usetikzlibrary{positioning}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue
}
\usepackage[nospace]{cite}

\input{macros}

%\title{Formal Analysis of EDHOC Key Establishment for Constrained IoT Devices}
\title{Analyzing the EDHOC Key Establishment Protocol using TAMARIN}
\author{
        Karl Norrman and 
        Vaishnavi Sundararajan and
        Alessandro Bruni
}

\begin{document}
\maketitle
%
\begin{abstract}
Given how common IoT devices that use constrained resources are becoming today, the need of the hour is communication protocols which can operate securely under such limitations.
%
For a few years, the Internet Engineering Task Force (IETF) has been working to standardize \mEdhoc{}, an authenticated key establishment protocol for such constrained IoT devices.
%
The first version of \mEdhoc{} was proposed in 2016.
%
In 2018, Bruni et al~\cite{DBLP:conf/secsr/BruniJPS18} used the \mProverif{} tool~\cite{DBLP:conf/csfw/Blanchet01} to formally analyze an early version of \mEdhoc{}, which had only two key establishment methods.  
%
%By now, the protocol has been fleshed out much more, and has been augmented with multiple new key establishment methods.
By 2021, the protocol had been fleshed out much more, with multiple new key establishment methods, and this version was formally analyzed using the \mTamarin{} prover~\cite{DBLP:conf/cav/MeierSCB13} in~\cite{Norr21}.
%
Here, we build on this work, and use \mTamarin{} to formally analyze the key establishment methods in the current version of \mEdhoc{}, as well as discuss some ramifications of the choices made while designing the protocol.
\vnote{We need to see what more we can say here, about the new stuff we're potentially adding}
%
%We show that not all methods satisfy the authentication notion
%injective agreement, but that all do satisfy a notion of implicit
%authentication, as well as Perfect Forward Secrecy (PFS) of the session key material.
%%
%We identify other weaknesses to which we propose improvements
%%
%For example, a party may intend to establish a session key with a certain peer,
%but ending up establishing it with another, trusted but compromised, peer.
%%
%We communicated our findings and proposals to the IETF, which
%has incorporated some of these in newer versions of the standard.
%Constrained IoT devices are becoming ubiquitous in society
%and there is a need for secure communication protocols that respect the
%constraints under which these devices operate.
%%
%\mEdhoc{} is an authenticated key establishment protocol for constrained IoT
%devices, currently being standardized by the Internet Engineering Task
%Force (IETF).
%%
%A rudimentary version of \mEdhoc{} with only two key establishment methods was
%formally analyzed in 2018.
%%
%Since then, the protocol has evolved significantly and several new key
%establishment methods have been added.
%%
%In this paper, we present a formal analysis of all \mEdhoc{} methods in an
%enhanced symbolic Dolev-Yao model using the \mTamarin{} tool.
%%
%We show that not all methods satisfy the authentication notion
%injective agreement, but that all do satisfy a notion of implicit
%authentication, as well as Perfect Forward Secrecy (PFS) of the session key material.
%%
%We identify other weaknesses to which we propose improvements
%%
%For example, a party may intend to establish a session key with a certain peer,
%but ending up establishing it with another, trusted but compromised, peer.
%%
%We communicated our findings and proposals to the IETF, which
%has incorporated some of these in newer versions of the standard.
\end{abstract}
%
%-------------------------------------------------------------------------- sec
\section{\uppercase{Introduction}}
\label{sec:introduction}
%%%%%%% START HERE %%%%%%%%%%
\input{intro}

%As IoT devices become more prevalent and get involved in progressively sensitive
%functions in society, the need to secure their communications
%becomes increasingly important.
%%
%Most security analysis has focused on computationally
%strong devices, such as cars and web-cameras, where existing protocols like
%\mDandTls{} suffice.
%%
%Constrained devices, on the other hand, which operate under severe
%bandwidth and energy consumption restrictions, have received much less
%attention.
%%
%These devices may be simple sensors, which only relay environment
%measurements to a server every hour, but need to function autonomously without
%maintenance for long periods of time.
%%
%The IETF standardized the Object Security for
%Constrained RESTful Environments (\mOscore{}) protocol to secure communications
%between constrained devices~\cite{rfc8613}.
%%
%However, the \mOscore{} protocol requires a pre-established security context.
%%
%The IETF has been discussing requirements and mechanisms for a key
%exchange protocol, named Ephemeral Diffie-Hellman Over COSE (\mEdhoc), for
%establishing \mOscore{} security contexts.
%%
%Naturally, \mEdhoc{} must work under the same constrained requirements as
%\mOscore{} itself.
%%
%While not all use cases for \mEdhoc{} are firmly set, the overall goal is to
%establish an \mOscore{} security context, under message size limitations.
%%
%It is therefore important to ensure that \mEdhoc{} satisfies fundamental
%security properties expected from a key exchange protocol.
%%The work is now being done in the IETF Lightweight Authenticated Key Exchange
%%(LAKE) working group.
%%
%
%The first incarnation of \mEdhoc{} appeared in March 2016.
%%
%It contained two different key establishment methods, one based on a
%pre-shared Diffie-Hellman (DH) cryptographic core%
%\footnote{By a \emph{cryptographic core}, or simply core, we mean an academic protocol,
%without encodings or application specific details required by an industrial
%protocol.
%%
%A \emph{key establishment method} is a core with some such details added.
%}
%and a second based on a
%variant of challenge-response signatures in the style of
%\mOptls{}~\cite{DBLP:conf/eurosp/KrawczykW16}.
%%
%\mEdhoc{} is therefore a framework of several key establishment methods.
%%
%In May 2018, the core based on challenge-response signatures was replaced by
%one based on \mSigma{} (SIGn-and-MAc)~\cite{sigma,bruni-analysis-selander-ace-cose-ecdhe-08}.
%%
%Since then the protocol has undergone significant changes.
%%
%Three new cores, mixing challenge-response signatures and regular signature for
%authentication, were added~\cite{our-analysis-selander-lake-edhoc-00}.
%%
%
%We formulate and formalize a security model covering all four key
%establishment methods, which is important especially since the
%\mSpec{}~\cite{our-analysis-selander-lake-edhoc-00} lacks a clear description
%of the intended security model and overall security goals.
%%
%
%We perform the analysis in a symbolic Dolev-Yao model.
%In this framework, we model messages as terms in an algebra, with operations such
%as encryption, modelled as functions on these terms.
%These functions are assumed perfect, e.g., one cannot decrypt an encrypted
%message without access to the key.
%The adversary, while unable to break encryption or reverse hashing, is modelled
%as the network.
%That is, the adversary, can block reroute, replay and modify messages at will.
%A symbolic model like this, while slightly severe an abstraction, still allows
%us to analyze \mEdhoc{} for logical flaws without incurring the complexity of
%a computational model.
%%
%The standardization process is ongoing, with the authors releasing newer
%versions of the \mSpec{} (see Section~\ref{sec:newdrafts} for more detail
%about how these versions differ from the one analyzed here).
%%
%
%%-------------------------------------------------------------------------- sub
%\subsection{Contributions}
%\label{sec:contributions}
%In this paper, we formally analyze the \mEdhoc{} protocol (with its four key
%establishment methods) using the \mTamarin{} tool~\cite{DBLP:conf/cav/MeierSCB13}.
%%
%We present a formal model we constructed of the protocol as given in the
%\mSpec{}~\cite{our-analysis-selander-lake-edhoc-00}.
%%
%
%We give an explicit adversary model for the protocol and verify
%properties such as session key material and entity authentication, and perfect
%forward secrecy, for all four methods.
%%
%
%The model itself is valuable as a basis for verifying further updates in the
%ongoing standardization.
%%
%It is publicly available~\cite{edhocTamarinRepo}.
%%
%It took several person-months to interpret the
%specification and construct the model.
%%
%Termination requires a hand-crafted proof oracle to guide \mTamarin{}.
%%
%
%We show that not all \mEdhoc{}'s key establishment
%methods provide authentication according to the injective agreement definition
%on the session key material, and none on the initiator's identity.
%%
%However, we show that all methods fulfill an implicit agreement property
%covering the session key material and the initiator's identity.
%%
%We identify a number of subtleties, ambiguities and weaknesses in the
%specification.
%%
%For example, the authentication policy requirements allows situations where a
%party establishes session key material with a trusted but compromised peer, even
%though the intention was to establish it with a different trusted party.
%%
%We provide remedies for the identified issues and have
%communicated these to the IETF and the specification authors, who incorporated
%some of our suggestions and currently consider how to deal with the remaining
%ones.
%%
%
%%-------------------------------------------------------------------------- sub
%\subsection{Comparison with Related Work}
%The May 2018 version of \mEdhoc{} was formally analyzed by
%\cite{DBLP:conf/secsr/BruniJPS18} using the \mProverif{}
%tool~\cite{DBLP:conf/csfw/Blanchet01}.
%%
%Their analysis covered a pre-shared key authenticated core and one
%based on \mSigma.
%%
%The properties checked for therein were secrecy, PFS and integrity of
%application data, identity protection against an active adversary,
%and strong authentication.
%%
%
%In contrast to the key establishment methods analyzed by Bruni et~al., which
%were based on the well-understood pre-shared key DH and \mSigma{} protocols,
%the three newly added
%methods combine two unilateral authentication protocols with the goal to
%constructing mutual authentication protocols.
%%
%Combining two protocols, which individually provide unilateral authentication,
%is not guaranteed to result in a secure mutual authentication
%protocol~\cite{DBLP:conf/ccs/Krawczyk16}.
%%
%Consequently, even though the framework is similar to the one analyzed by Bruni
%et~al., the cryptographic underpinnings have significantly increased in
%complexity, and is using mechanisms which have not previously been formally analyzed.
%%
%The set of properties we check for is also different.
%%
%Our analysis is further carried out using a different tool,
%namely \mTamarin; different kinds of strategies to formulate and
%successfully analyze the protocol are required when working with this tool.
%

%-------------------------------------------------------------------------- sec
\section{\uppercase{The \mEdhoc{} Protocol}}
\label{sec:edhoc}
\input{protocol}

%-------------------------------------------------------------------------- sec
\section{\uppercase{Formalization and Results}}
\label{sec:formalization}
\input{edhocFormalization}

%-------------------------------------------------------------------------- sec
\section{\uppercase{Denial of Service}}
\label{sec:dos}
\mPoint{DoS is out of scope for \mEdhoc{}; apps to deal with it}
\mEdhoc{} includes no specific measures for countering Denial of Service (DoS)
attacks and delegates the responsibility to handle this to the application.
%
\mEdhoc{} itself has limited functionality, few negotiable parameters and
processing rules
%
Hence, we argue in this section that the attack surface
is small from a DoS perspective.
%
However, \mEdhoc{} leaves much of the semantics of its messages undefined.
%
This increases the risk of applications vulnerable to DoS even if \mEdhoc{} is
arguably not vulnerable to DoS itself.
%

\mPoint{\mEdhoc{} gives a little info of questionable use for developers}
Section 8.7 of the \mEdhoc{} specification explains the design philosophy w.r.t.
DoS in two paragraphs and gives two recommendations.
%
The first being that mitigation of DoS based on sending proper messages should
be handled by lower layers, e.g., by checking reachability of the sending
party\footnote{Karl: this may be another way to perform reflection DoS attacks}.
%
The second being that an application may try to determine (in some way) that an
apparently valid message is in fact probably a forgery and should be ignored or
cause processing to stop.
%

\mPoint{Responder is more vulnerable than initiator}.
For the same reasons as given in the IKEv2 DDoS analysis
by Nir and Smyslov~\cite{rfc8019} the responder is more vulnerable than the
initiator.
%
From a roles and message passing perspective, IKEv2 is very similar to all
methods in \mEdhoc{}.
%
The RFC provides much good advice for protocols with this structure that
\mEdhoc{} using applications can reuse.
%
We should look more into this.
%

\mPoint{Amplification}
An adversary can fill a message 1 with easily generated bogus data and send this
to a victim responder.
%
There is little work for the initiator, but both symmetric and asymmetric key
operations required by the responder.
%
\mEdhoc{} does not protect against this, but refer to the application and its
transport layer to resolve the problem.
%
This is a design decision, see Section 8.7 of \mEdhoc{} specification.
%
In addition to the computational resources the responder spends, it may also
need to obtain certificate chains and verify certificate revocation servers as a
result of the spoofed message.
%
The asymmetry between the adversaries effort and the responder may hence be
significant.
%
Since \mEdhoc{} specifically targets constrained devices this asymmetry needs to
be considered even more carefully than for IKEv2~\cite{rfc8019}.
%

Recommend that the role of responder is chosen carefully.
%
The responder is most vulnerable.
%
For example, is it better that a lock or one of the key cards takes on the
responder role and becomes vulnerable?
%

\mPoint{Reflection Attacks}
Reflection attacks are attacks where an adversary spoofs the source address of a
victim in a message and thereby tricks the responder to send a message to the
victim.
%
\mEdhoc{} leaves protection against this to the transport protocol and
application, and suggests using a path reachability method similar to IKEv2.
%
In situations where the responder does not need to perform additional
communications to verify certificates or similar, a reachability mechanism may
be more expensive than performing the computations.
\footnote{Find ref of that computation is often cheaper than radio
transmission, at least energy-wise.}
%
However, if round-trip times are large, and half-open states hence are kept
by the receiver for a long time, a reachability mechanism can trade
communication for storage.
\footnote{RFC8019 discuss some numbers as examples of tradeoffs. Should we set
up a simple model and derive an expression for trade-offs? Could be useful for
implementors to consider, but low academic value. This would capture also
amplification.}
%

\mPoint{In a constrained device randomness may be easily depleted}
Assume an adversary generate (deterministic) bogus message 1s and send these at
a sufficiently high rate to the responder.
%
\mEdhoc{} specifically targets constrained devices, that presumably may not
always have good randomness sources.
%
Is it possible to deplete or significantly reduce the entropy of the source and
in this way be able to predict the responder's random input to other sessions
(with other initiators)?
%
The LAKE WG have been talking about effects of poor randomness on signatures, so
this should be an issue, no?
%

\mPoint{\mEdhoc{} can deadlock if not used correctly}
An adversary may send forged error messages, and specifically reject proposed
ciphersuites.
%
The initiator is recommended to not try that ciphersuite with the responder
again in a new session for a while.
%
If the adversary rejects all options for a initiator this way, it can prevent
communication between the initiator and that responder until the cache is
cleared.
%
Potentially, an adversary could lock a party of a group out of communication all
together this way.
%
This needs to be considered by applications.
%

\mPoint{Is any of the methods better than the others from a DoS perspective?}
We considered this for quite some time, but there seems to be little difference
in this regard.
%

%-------------------------------------------------------------------------- sec
\section{\uppercase{Discussion}}
\label{sec:discussion}
There are a few places where \mEdhoc{} can be improved,
which we found during this work and communicated to the authors.
%
We discuss them below.
%

%-------------------------------------------------------------------------- sub
\subsection{Unclear Intended Use}
\label{sec:unclearProtocolUse}
%
The \mEdhoc{} \mSpec{} lists several security goals, but they are
imprecise and difficult to interpret due to lack of context and intended usage
descriptions.
%
Without knowing how the protocol is to be used,
it is not clear whether the listed security goals are the most important ones
for constrained IoT devices.
%

The abstract goal of \mEdhoc{} is simple: establish an \mOscore{} security
context using few roundtrips and small messages.
%
From that, the design of \mEdhoc{} is mainly driven by what
can be achieved given the technical restrictions.
%
Focusing too much on what can be achieved within given restrictions, and paying
too little attention to the use cases where the
protocol is to be used and their specific goals, risks resulting in
sub-optimal trade-offs and design decisions.
%

\mEdhoc{} is intended to cover a variety of use cases, many of which are
difficult to predict today.
%
However, this does not
prevent collecting \emph{typical} use cases and user stories
to identify more specific security goals that will be important in most cases.
%

While constructing our model, we made up simple user stories to identify
security properties of interest.
%
Several of these revealed subtleties and undefined aspects of \mEdhoc{}.
%
We informed the \mEdhoc{} authors, who addressed these aspects in the
\mSpec{}.
%

\subsubsection{(Non-)Repudiation}
An access control solution for a nuclear power-plant may need to log who is
passing through a door, whereas it may be undesirable for, say, a coffee
machine to log a list of users along with their coffee preferences.
%
Via this simple thought experiment, we realized that the \mSpec{} did not
consider the concept of (non)-repudiation.
%
In response, the authors of the \mSpec{} added a paragraph discussing how
different methods relate to (non)-repudiation.

\subsubsection{Unintended Peer Authentication}
Section~3.2 of the \mSpec{} states that parties must be configured
with a policy restricting the set of peers they run \mEdhoc{} with.
%
However, the initiator is not required to verify that the \mIdcredr{} received
in the second message is the same as the one intended.
%
The following attack scenario is therefore possible.
%

Suppose someone has configured all devices in their home to be in the
allowed set of devices, but that one of the devices ($A$) is compromised.
%
If another device $B$, initiates a connection to a third device $C$, the
compromised device $A$ may interfere by responding in $C$'s place, blocking
the legitimate response from $C$.
%
Since $B$ does not verify that the identity indicated in the second message
matches the intended identity $C$, and device $A$ is part of the allowed set,
$B$ will complete and accept the \mEdhoc{} run with device $A$ instead of the
intended $C$.
%
The obvious solution is for the initiator to match \mIdcredr{} to the intended
identity indicated by the application, which we included in our model.
%
We have communicated this to the \mEdhoc{} authors and they are considering
how to resolve the issue.
%

%------------------------------------------------------------------------- sub
\subsection{Unclear Security Model}
We argue that the \mSpec{} gives too little information about what capabilities
an adversary is assumed to have, and that this leads to unclear design goals and
potentially sub-optimal design.
%

Even though \mEdhoc{} incorporates cryptographic cores from different academic
security protocols, its design does not take into account the adversary models
for which these protocols were designed.
%
For example, \mOptls{}, whose cryptographic core is essentially the same
as the \mStat{} authentication method, is designed to be secure in the CK
model~\cite{DBLP:conf/crypto/CanettiK02}.
%
The CK security model explicitly separates the secure storage of long-term
keys from storage of session state and ephemeral keys.
%
This is appropriate for modelling the use of secure modules.
%

The \mEdhoc{} authors indicated to us that it was
not necessary to consider compromised ephemeral keys separately from
compromised long-term keys.
%
The rationale is that \mSigma{} cannot protect against compromised ephemeral
keys~\cite{personalCommunication}.
%
That rationale is presumably based on the fact that the \mSigSig{} method is
closely modeled on the \mSigmaI{} variant of \mSigma{}, and that it would be
preferable to obtain a homogeneous security level among the \mEdhoc{}
methods.
%
That rationale is only true, however, if one restricts attention to session key
confidentiality of an ongoing session.
%
Secure modules provide value in other ways, for example, by allowing
constructions with Post-Compromise Security (PCS) guarantees.
%
We discussed this with the authors, and
the latest version of the \mSpec{}~\cite{latest-ietf-lake-edhoc-05} includes
recommendations on storage of long-term keys and operations on these inside a
secure module.
%

%-------------------------------------------------------------------------- sub
\subsection{Session Key Material}
\label{sec:sessionKeyMaterial}
\mEdhoc{} establishes session key material, from which session keys
can be derived using the \mEdhoc{}-Exporter.
%
The session key material is affected by \mGxy{}, and if a party uses the
\mStat{} authentication method, also by that party's secret static long-term key.
%
As shown in Section~\ref{sec:formalization}, mutual injective agreement cannot
be achieved for $P_I$.
%
If this property is not important for constrained IoT devices which cannot use
any of the other methods, then one can simply accept that the methods have
different authentication strengths.
%
Otherwise, this is a problem.
%

We identified three alternatives for resolving this.
%
One alternative is to include \mIdcredi{}, or its hash, in the first and
second messages.
%
This would, however, increase message sizes and prevent initiator identity
protection, which are grave concerns for \mEdhoc{}.
%
A second alternative is to not derive the session key material from $P_I$.
%
Doing so, however, deviates from the design of \mOptls{} (and similar protocols
from which the \mStat{}-based methods are derived), where the inclusion of
$P_I$ plays a crucial part in the security proof of resistance against
initiator ephemeral key compromise.
%
The third alternative is to include a fourth message from responder to initiator,
carrying a MAC based on a key derived from session key material including $P_I$.
%
Successful MAC verification guarantees
to the initiator that the responder injectively agrees on $P_I$.
%
We presented the options to IETF, and they decided to add a
fourth message as an option in the latest version of the
\mSpec{}~\cite{latest-ietf-lake-edhoc-05}.
%

Regardless of how this is handled, we verified that all methods
enjoy a common, but weaker, property: mutual implicit agreement
on all of $P_e, P_I$ and $P_R$, where applicable.
%

%-------------------------------------------------------------------------- sec
\section{\uppercase{Conclusions and Future Work}}
\label{sec:conclusions}
\label{sec:newdrafts}
We formally modeled all four
methods of the \mEdhoc{} \mSpec{} using \mTamarin.
%
We formulated several important security properties and identified precise
adversary models in which we verified these.
%
The properties are shown in Table~\ref{tab:props}.
%
Mutual injective agreement covers the parameters $S_P$:
responder identity, roles, session key material (except for $P_I$ when
initiator uses the \mStat{} authentication
method), context identifiers \mCi{} and \mCr, and cipher suites \mSuites.
%
The responder in addition is ensured agreement on the initiators identity and
$P_I$, i.e., on the set $S_F$.
%
Implicit agreement covers all previously mentioned parameters for both peers.
%
Verification of all lemmas, including model validation lemmas, took 42 minutes
on an Intel Core i7-6500U 2.5GHz using two cores.
%
Mutual entity authentication, UKS- and KCI resistance can be inferred
from the verified properties.
%
\begin{table*}[h!]
        \centering
        \caption{Verified properties. $S_P$ contains
            roles, responder identity, session key material (excluding
            $P_I$), \mCi, \mCr, and \mSuites. $S_F$ is $S_{P}$,
            the initiator identity, and $P_I$.}
        \label{tab:props}
        \begin{tabular}{|l|c|c|c|c|}
                \hline
                & \mSigSig & \mSigStat & \mStatSig & \mStatStat \\
                \hline
                Injective agreement for I & $S_F$ & $S_F$ & $S_P$ & $S_P$\\
                Injective agreement for R & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
                Implicit agreement for I & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
                Implicit agreement for R & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
                PFS for session key material & \cm & \cm & \cm & \cm\\
                \hline
        \end{tabular}
\end{table*}

Further, we identified a situation where initiators may establish an \mOscore{}
security context with a different party than the application using \mEdhoc{}
intended, and proposed a simple mitigation.
%
We discussed how the IETF may extract and better define security properties to
enable easier verification.

We verified each method in isolation, and leave as future work to verify whether
the methods are secure under composition.

%\subsection{Newer Versions of the Specification}
%\label{sec:newdrafts}
In this work, we have analyzed the \mEdhoc{} version as of July
2020~\cite{our-analysis-selander-lake-edhoc-00}.
%
There are newer versions, with the most recent version as
of February 2021~\cite{latest-ietf-lake-edhoc-05}.
%
However, the changes to the protocol over these versions are not
particularly significant for our analysis.
%

%-------------------------------------------------------------------------- ack
\section*{ACKNOWLEDGEMENTS}
This work was partially supported by
the Wallenberg AI, Autonomous Systems and Software Program (WASP) funded by
the Knut and Alice Wallenberg Foundation.
%
We are grateful to G\"oran Selander, John Mattsson and Francesca Palombini for
clarifications regarding the specification.
%

%-------------------------------------------------------------------------- bib
\bibliographystyle{apalike}
{\small
    \bibliography{refComp}
}
\end{document}