### 2022-07-12 Journal Version/protocol.tex

% !TEX root = journal-version.tex
In this section, we describe the various key establishment methods of the \mEdhoc{} protocol.
%
Following~\cite{Norr21}, we refer to the two roles executing the protocol as the initiator $I$ and the responder $R$.
%
We annotate values with $I$ and $R$ to make explicit which role they belong to.
%

We denote by $\langle d_{t, \mathit{id}}, Q_{t, \mathit{id}}\rangle$ public-private key pairs, where $d$ is the private key, $Q$ is the public key, $t \in \{e, s\}$ indicates whether the key is ephemeral or static, and $\mathit{id}$ indicates the party to whom the key pair belongs.
%
When clear from context, we will often drop part (or all) of the subscript.
%
Static key pairs (suitable for both regular and challenge-response signatures) are long-term authentication credentials, while ephemeral key pairs are those which are generated afresh for each instantiation of the protocol.
%

Parties can authenticate using regular signatures or challenge-response signatures. 
%
In the former case, we say that they use the \emph{signature based authentication method} (\mSig{}).
%
In the latter, we say, following the terminology in the \mSpec{}, that they use the \emph{static key authentication method} (\mStat{}).
%
For more details about the challenge-response terminology, see~\cite{DBLP:conf/crypto/Krawczyk05}.

%%------------------------------------------------------------------------- sub
%

\mEdhoc{} relies on \mCose{}~\cite{rfc8152} for elliptic curve operations and
transforming points into bitstrings, and we therefore abstract those as
follows.
%
Signatures and verification thereof using party $A$'s key pair are
denoted \mSign{A}$(\cdot)$ and \mVf{A}$(\cdot)$ respectively.
%
The DH-primitive combining a private key $d$ and a point $P$ is denoted
$\mDH(d,P)$.
%
We abuse notation and let these function symbols denote operations on both
points and the corresponding bitstrings.
%

%------------------------------------------------------------------------- sub
\subsection{Framework Structure}
\label{sec:framework}
\mEdhoc{}'s goal is to establish an \mOscore{} security context,
including session key material \mbox{denoted \mSessKey{}}, and optionally transfer
application data \mADone{}, \mADtwo{} and \mADthree{}.
%
To accomplish this, the
\mSpec{}~\cite{our-analysis-selander-lake-edhoc-00} gives
a three-message protocol pattern, shown in Figure~\ref{fig:edhocFramework}.
%
We first describe this pattern and the parts that are common to all key
establishment methods.
%
Then we describe authentication and derivation of keys in more detail.
%
The latter is what differentiate the key establishment methods from each other.
%

\subsubsection{Protocol Pattern}
The first two messages negotiate authentication methods \mMethod{} and a
ciphersuite \mSuites{}.
%
In \mMethod{}, the initiator $I$ proposes which authentication method each
party should use.
%
These may differ, leading to four possible combinations:
\mSigSig{}, \mSigStat{}, \mStatSig{} and \mStatStat{}.
%
We refer to these \emph{combinations} of authentication methods simply as
\emph{methods} to align with the \mSpec{} terminology.
%
The first authentication method in a combination is the one proposed for the
initiator and the last is the one proposed for the responder.
%
The responder $R$ may reject the choice of method
or cipher suite with an error message, resulting in negotiation across
multiple \mEdhoc{} sessions.
%
Our analysis excludes error messages.
%

\mEdhoc{}'s first two messages also exchange connection identifiers \mCi{} and
\mCr{}, and public ephemeral keys, \mGx{} and \mGy{}.
%
The connection identifiers \mCi{} and \mCr{}, described in Section 3.1 of the
\mSpec{}, deserve some elaboration.
%
The \mSpec{} describes these identifiers as not serving a security purpose for
\mEdhoc{}, but only as aiding message routing to the correct \mEdhoc{} processing
entity at a party.
%
Despite this, the \mSpec{} states that they may be used by \mOscore{}, or other
protocols using the established security context, without restricting how they
are to be used.
%
Because \mEdhoc{} may need them in clear-text for routing, \mOscore{} cannot
rely on them being secret.
%
Section 7.1.1 of the \mSpec{} requires the identifiers to be unique.
%
Uniqueness is defined to mean that $\mCi{} \not = \mCr{}$ for a given session
and the \mSpec{} requires parties to verify that this is the case.
%
The same section also require that \mOscore{} must be able to retrieve the
security context based on these identifiers.
%
The intended usage of \mCi{} and \mCr{} by \mOscore{} is not made specific and
therefore it is not clear which properties should be verified.
%
We verify that the parties agree on the established values.
%

The two last messages provide identification and authentication.
%
Parties exchange identifiers for their long-term keys, \mIdcredi{} and \mIdcredr{},
as well as information elements, \mAuthi{} and \mAuthr{}, to authenticate
that the parties control the corresponding long-term keys.
%
The content of \mAuthi{} and \mAuthr{} depends on the authentication method
associated with the corresponding long-term key.
%
For example, if $\mMethod{} = \mSigStat{}$, the responder $R$ must either
reject the offer or provide an \mIdcredr{} corresponding to a key pair
$\langle\mPriv{R},\ \mPub{R}\rangle$ suitable for use with challenge-response
signatures, and
compute \mAuthr{} based on the static key authentication method \mStat{}.
%
In turn, the initiator $I$ must provide an \mIdcredi{} corresponding to a key
pair $\langle\mPriv{I},\ \mPub{I}\rangle$ suitable for a regular signature,
and compute
\mAuthi{} based on the signature based authentication method \mSig{}.
%

\subsubsection{Authentication}
Regardless of whether \mStat{} or \mSig{} is used to compute \mAuthr{}, a
MAC is first computed over \mIdcredr{}, \mCredr{}, a transcript hash of the
information exchanged so far, and \mADtwo{} if included.
%
The MAC is the result of encrypting the empty string with the Authenticated
Encryption with Additional Data (AEAD) algorithm from the ciphersuite
\mSuites{}, using the mentioned information as additional data.
%
The MAC key is derived from the ephemeral key material
\mGx{}, \mGy{}, \mX{} and \mY{}, where $I$
computes $\mDH(\mX,\ \mGy)$ and $R$ computes $\mDH(\mY,\ \mGx)$, both resulting in
the same output in the usual DH way.
%

In case $R$ uses the \mSig{} authentication method, \mAuthr{} is $R$'s
signature over the MAC itself and the data that the MAC already covers.
%
In case $R$ uses the \mStat{} authentication method, \mAuthr{} is simply the
MAC itself.
%
However, when using \mStat{}, the key for the MAC is derived, not only from the
ephemeral key material, but also from $R$'s long-term key
$\langle\mPriv{R},\ \mPub{R}\rangle$.
%
For those familiar with \mOptls, this corresponds to the 1-RTT semi-static
pattern computing the MAC key \textsf{sfk} for the \textsf{sfin}
message~\cite{DBLP:conf/eurosp/KrawczykW16}.
%
The content of \mAuthi{} is computed in the corresponding way for the initiator
$I$.
%
In Figure~\ref{fig:edhocFramework}, we denote a MAC using a key derived from
both $\langle\mPriv{R},\ \mPub{R}\rangle$ and $\langle\mX,\ \mGx\rangle$ by
$\mathit{MAC}_I$, and a MAC using a key derived from
both $\langle\mPriv{I},\ \mPub{I}\rangle$ and $\langle\mY,\ \mGy\rangle$ by
$\mathit{MAC}_R$.
%

Parts of the last two messages are encrypted and integrity protected, as
indicated in~\ref{fig:edhocFramework}.
%
The second message is encrypted by XORing the output of the key derivation
function HKDF (see Section~\ref{sec:keysched}) on to the plain text.
%
The third message is encrypted and integrity protected by the AEAD algorithm
determined by the ciphersuite \mSuites{}.
%

\begin{figure}
\centering
\tikzset{>=latex, every msg/.style={draw=thick}, every node/.style={fill=none,text=black}}
\begin{tikzpicture}
    \node (ini) at (0, 0) {Initiator};
    \draw [very thick] (0, -0.25) -- (0,-2.3);
    \draw [very thick] (5.75, -0.25) -- (5.75,-2.3);
    \node (res) at (5.75,0) {Responder};
    \msg{1em}{ini}{res}{\mMsgone: \mMethod, \mSuites, \mGx, \mCi, \mADone};
    \msg{3em}{res}{ini}{\mMsgtwo: \mCi, \mGy, \mCr, \{\mIdcredr, \mAuthr, \mADtwo\}};
    \msg{5em}{ini}{res}{\mMsgthree: \mCr, \{\mIdcredi, \mAuthi, \mADthree\}};
    \draw [line width=1mm] (-0.75,-2.3) -- (0.75,-2.3);
    \draw [line width=1mm] (5.75-0.75,-2.3) -- (5.75+0.75,-2.3);
    \node (padding) at (0,-2.5) {};
    \end{tikzpicture}
    \begin{tabular}{|c|c|c|}
        \hline
        \mMethod & \mAuthi & \mAuthr\\
        \hline
        \mSigSig{} & $\mSign{I}(\cdot)$ & $\mSign{R}(\cdot)$ \\
        \mSigStat{} & $\mSign{I}(\cdot)$ & $\textit{MAC}_R(\cdot)$\\
        \mStatSig{} & $\textit{MAC}_I(\cdot)$ & $\mSign{R}(\cdot)$\\
        \mStatStat{} & $\textit{MAC}_I(\cdot)$ & $\textit{MAC}_R(\cdot)$\\
        \hline
    \end{tabular}
    \caption{Structure of \mEdhoc{}: $\{t\}$ means $t$ is encrypted and integrity
protected.}
\label{fig:edhocFramework}
\end{figure}

\subsubsection{Key Schedule}
\label{sec:keysched}
At the heart of \mEdhoc{} is the key schedule depicted in
Figure~\ref{fig:kdfdiagram}.
%
\mEdhoc{} uses two functions from the \mHkdf{} interface~\cite{rfc5869} to derive keys.
%
\mHkdfExtract{} 
constructs uniformly distributed key material from random input and a salt,
while \mHkdfExpand{} generates keys from key material and a salt.
%

The key schedule is rooted in the ephemeral DH key
\mGxy{}, which is computed as $\mDH(\mX, \mGy)$ by $I$ and as $\mDH(\mY, \mGx)$
by $R$.
%
From \mGxy{}, three intermediate keys \mPRKtwo, \mPRKthree{} and
\mPRKfour{} are derived during the course of protocol execution.
%
Each of them is used for a specific message in the protocol, and from
these intermediate keys, encryption and integrity keys
(\mKtwoe, \mKtwom{}, \mKthreeae, and \mKthreem) for that message are derived.
%
The salt for generating \mPRKtwo{} is the empty string.
%

The protocol uses a running transcript hash $th$, which includes the information
transmitted so far.
%
The value of the hash, denoted $th_i$ for the $i$th message, is included in key
derivations as shown in Figure~\ref{fig:kdfdiagram}.
%

Successful protocol execution establishes the session key material \mSessKey{}
for \mOscore{}.
%
\mSessKey{} can be considered a set that always includes \mGxy{}.
%
If the initiator uses the \mStat{} authentication method, \mSessKey{} also
includes
$\mDH(\mY,\ \mGi{}) = \mDH(\mPriv{I},\ \mGy)$, which we denote $P_I$.
%
If the responder uses the \mStat{} authentication method, it also includes
$\mDH(\mX,\ \mGr{}) = \mDH(\mPriv{R},\ \mGx)$, which we denote $P_R$.
%
From the session key material, a key exporter (\mEdhoc-Exporter) based on
\mHkdf{} is used to extract keys required for \mOscore{}.
%

\begin{figure*}[!h]
\centering
\scalebox{.785}{
\input{kdfdiagram.tex}
}
\caption{Key schedule: Light blue boxes hold DH keys ($P_e, P_I, P_R$),
orange boxes intermediate key material (\mPRKtwo, \mPRKthree, \mPRKfour), and
dark blue boxes keys for \mAead{} or \mXor{} encryption
(\mKtwoe, \mKtwom, \mKthreeae, \mKthreem).
Dashed boxes are conditionals.}
\label{fig:kdfdiagram}
\end{figure*}

As an illustrative example of the entire process, we refer to
Figure~\ref{fig:edhocsigstat}, which depicts the protocol pattern, operations
and key derivations for the \mSigStat{} method in more detail.
%
\begin{figure*}[ht]
\centering
\scalebox{.7}{
\tikzset{>=latex, every msg/.style={draw=thick}, every node/.style={fill=none,text=black}}
\begin{tikzpicture}
    \node (ini) at (0, 0) {Initiator};
    \draw [very thick] (0, -0.5) -- (0,-14.8);
    \draw [very thick] (9, -0.5) -- (9,-14.8);
    \node[below=0.5em of ini,fill=white] {$
    \begin{array}{c}
        \text{Knows}\ \langle\mPriv{I},\ \mPub{I}\rangle,\ \mIdcredi,\ \mIdcredr,\ \mADone,\ \mADthree
    \end{array}
    $};
    \node (res) at (9,0) {Responder};
    \node[below=0.5em of res,fill=white] {$
    \begin{array}{c}
        \text{Knows}\ \langle\mPriv{R},\ \mPub{R}\rangle, \ \mIdcredr,\ \mADtwo
    \end{array}$};
    \action{3.5em}{ini}{Generates $\mMethod,\ \mSuites,\ \mCi,\ \langle\mX{},\ \mGx\rangle$};
    \msg{6.5em}{ini}{res}{\mMsgone: \mMethod, \mSuites, \mGx, \mCi, \mADone};
    \action{7.0em}{res}{$
      \begin{array}{c}
          \text{Generates } \mCr,\ \langle\mY{},\ \mGy\rangle\\
          \ \ P_e = \mDH(\mY,\ \mGx{})\\
          \ \ P_R = \mDH(\mPriv{R},\ \mGx{})\\
        \mTHtwo = \mHash(\mMsgone, \langle \mCi, \mGy, \mCr \rangle)\\
        \mPRKtwo = \mHkdfExtract(\textrm{``\phantom{}''}, P_e) \\
        \mPRKthree = \mHkdfExtract(\mPRKtwo, P_R) \\
        \mKtwom = \mHkdfExpand(\mPRKthree, \mTHtwo) \\
        \mMactwo = \mAead(\mKtwom; \langle \mIdcredr, \mTHtwo, \mCredr, \mADtwo \rangle; \textrm{``\phantom{}''}) \\
        \mKtwoe = \mHkdfExpand(\mPRKtwo, \mTHtwo)
      \end{array}$};
    \msg{21.7em}{res}{ini}{\mMsgtwo: \mCi, \mGy, \mCr, $\overbrace{\mKtwoe\ \mXor\ \langle \mIdcredr, \mMactwo, \mADtwo \rangle}^{\mCipher}$};
    \action{22.5em}{ini}{$
      \begin{array}{c}
        %\mTHtwo = \mHash(\mMsgone, \langle \mCi, \mGy, \mCr \rangle) \
        \ P_e = \mDH(\mX,\ \mGy{})\\
        \mPRKtwo = \mHkdfExtract(\textrm{``\phantom{}''}, P_e) \\
        %\mKtwoe = \mHkdfExpand(\mPRKtwo,\mTHtwo)\\
        %\mGrx = \mCredr^{x} \\
        \ \ P_R = \mDH(\mX,\ \mPub{R})\\
        \mPRKfour = \mPRKthree = \mHkdfExtract(\mPRKtwo, P_R) \\
        %\mKtwom = \mHkdfExpand(\mPRKthree, \mTHtwo) \\
        \mKthreeae = \mHkdfExpand(\mPRKthree, \mTHtwo) \\
        \mTHthree = \mHash(\mTHtwo, \mCipher, \mCr)\\
        \mKthreem = \mHkdfExpand(\mPRKfour, \mTHthree) \\
        \mMacthree = \mAead(\mKthreem; \langle \mIdcredi, \mTHthree, \mCredi, \mADthree \rangle; \textrm{``\phantom{}''}) \\
        \mSigthree = \mSign{I}(\langle \langle \mIdcredi, \mTHthree, \mCredi, \mADthree \rangle, \mMacthree \rangle)
      \end{array}$};
    \msg{35.5em}{ini}{res}{$\mMsgthree: \mCr, \mAead(\mKthreeae; \mTHthree; \langle \mIdcredi, \mSigthree, \mADthree \rangle$)};
    \action{36em}{res}{$
    \begin{array}{c}
        \mTHthree = \mHash(\mTHtwo, \mCipher, \mCr)\\
        \mKthreem = \mHkdfExpand(\mPRKthree, \mTHthree) \\
        \mKthreeae = \mHkdfExpand(\mPRKthree, \mTHthree)
    \end{array}$};
    \draw [line width=1mm] (-2,-14.8) -- (2,-14.8);
    \draw [line width=1mm] (7,-14.8) -- (11,-14.8);
    \end{tikzpicture}
}
    \caption{The \mSigStat{} method.
    Tuples are denoted $\langle\cdot\rangle$, and the hash function \mHash{} is
    as determined by \mSuites{}.}
\label{fig:edhocsigstat}
\end{figure*}
%------------------------------------------------------------------------- sub