### 2022-07-15 protocol.tex

% !TEX root = paper.tex
In this section, we describe the various key establishment methods of the \mEdhoc{} protocol.
%
Following~\cite{Norr21}, we refer to the two roles executing the protocol as the initiator $I$ and the responder $R$.
%
We annotate values with $I$ and $R$ to make explicit which role they belong to.
%

%------------------------------------------------------------------------- sub
\subsection{Notation}
\label{sec:notation}
We denote by $\langle d_{t, \mathit{id}}, Q_{t, \mathit{id}}\rangle$ public-private key pairs, where $d$ is the private key, $Q$ is the public key, $t \in \{e, s\}$ indicates whether the key is ephemeral or static, and $\mathit{id}$ indicates the party to whom the key pair belongs.
%
When clear from context, we will often drop part (or all) of the subscript.
%
Static key pairs (suitable for both regular and challenge-response signatures) are long-term authentication credentials, while ephemeral key pairs are those which are generated afresh for each instantiation of the protocol.
%

Parties can authenticate using regular signatures or challenge-response signatures. 
%
In the former case, we say that they use the \emph{signature based authentication method} (\mSig{}).
%
In the latter, we say, following the terminology in the \mSpec{}, that they use the \emph{static key authentication method} (\mStat{}).
%
For more details about the challenge-response terminology, see~\cite{DBLP:conf/crypto/Krawczyk05}.
%

\mEdhoc{} fundamentally utilizes elliptic curves and associated Diffie-Hellman operations. 
%
Signatures using a principal $A$'s keys are denoted by \mSign{A}$(\cdot)$, while the verification thereof is denoted by \mVf{A}$(\cdot)$.
%
A Diffie-Hellman operation which combines a private key $d$ and a point $P$ on the elliptic curve is represented as $\mDH(d, P)$.
%
We will often overload notation to let $P$ stand for both the point on the elliptic curve as well as the corresponding bitstring.

%\mEdhoc{} relies on \mCose{}~\cite{rfc8152} for elliptic curve operations and
%transforming points into bitstrings, and we therefore abstract those as
%follows.
%
%Signatures and verification thereof using party $A$'s key pair are
%denoted \mSign{A}$(\cdot)$ and \mVf{A}$(\cdot)$ respectively.
%%
%The DH-primitive combining a private key $d$ and a point $P$ is denoted
%$\mDH(d,P)$.
%%
%We abuse notation and let these function symbols denote operations on both
%points and the corresponding bitstrings.
%

%------------------------------------------------------------------------- sub
\subsection{Overall description}
\label{sec:description}
\mEdhoc{} is designed to establish a security context for \mOscore{}.
%
This context includes the session key material (we denote this by \mSessKey{}), and might also transfer application data \mADone{}, \mADtwo{}, and \mADthree{}.
%
The abstract protocol as given in the \mSpec{}~\cite{} consists of three messages, and is shown in Figure~\ref{fig:edhocFramework}~\cite{Norr21}.
%
The generalized abstract protocol is the same across methods, but the authentication mechanisms and key derivation procedures differ. 
%
We will describe the various aspects of the protocol now.

\begin{figure}[ht]
\centering
\tikzset{>=latex, every msg/.style={draw=thick}, every node/.style={fill=none,text=black}}
\begin{tikzpicture}
    \node (ini) at (0, 0) {Initiator};
    \draw [very thick] (0, -0.25) -- (0,-2.3);
    \draw [very thick] (5.75, -0.25) -- (5.75,-2.3);
    \node (res) at (5.75,0) {Responder};
    \msg{1em}{ini}{res}{\mMsgone: \mMethod, \mSuites, \mGx, \mCi, \mADone};
    \msg{3em}{res}{ini}{\mMsgtwo: \mCi, \mGy, \mCr, \{\mIdcredr, \mAuthr, \mADtwo\}};
    \msg{5em}{ini}{res}{\mMsgthree: \mCr, \{\mIdcredi, \mAuthi, \mADthree\}};
    \draw [line width=1mm] (-0.75,-2.3) -- (0.75,-2.3);
    \draw [line width=1mm] (5.75-0.75,-2.3) -- (5.75+0.75,-2.3);
    \node (padding) at (0,-2.5) {};
    \end{tikzpicture}
    \caption{Structure of \mEdhoc{}: $\{t\}$ means $t$ is encrypted and integrity
protected.~\cite{Norr21}}
\label{fig:edhocFramework}
\end{figure}

Of the three messages, the first two, among other things, serve to establish a common authentication method \mMethod{}, and a ciphersuite \mSuites{}.
%
In \mMethod{}, the party playing the Initiator role proposes an ordered list of choices for the authentication methods that each party can use. 
%
As mentioned earlier, these may differ for the two roles, yielding four possible combinations: \mSigSig{}, \mSigStat{}, \mStatSig{}, and \mStatStat{}, where the first is the method used by the Initiator role, and the second by the Responder role.\footnote{As in the \mSpec{}, we will overload notation and refer to the combinations of authentication methods as methods as well.}
%
The party executing the Responder role may choose to reject the method or ciphersuite chosen by the Initiator by sending an error message.
%
This results in abandoning this session and renegotiating, as the Initiator goes down their list of choices for authentication methods, and picks the next option.
%
While our analysis (and indeed, \mTamarin{}) cannot naturally handle such renegotiation which maintains state and ``remembers'' the rejected method and/or ciphersuite, the formal modeling excludes this aspect of the protocol. 
%
However, we will discuss the ramifications of such a renegotiation procedure and the error messages later, in Section~\ref{sec:}.
%

In addition to negotiating the method and the ciphersuite, the first two messages are also instrumental for the exchange of public ephemeral keys \mGx{} and \mGy{}, and connection identifiers \mCi{} and \mCr{}, for the initiator and responder roles respectively. 
%
The \mSpec{} states that the connection identifiers serve only to route messages to the correct principal executing \mEdhoc{}, but then also claims that they may be used in turn by any protocols (like \mOscore{}) using the security context established by \mEdhoc{}. 
%
While the \mSpec{} does not require any explicit security guarantees to be satisfied by these communication identifiers, it does, however, require that the identifiers be unique, i.e. in any session, $\mCi{} \neq \mCr{}$, and that the parties involved in the session can verify this uniqueness.
%
The \mSpec{} states that \mOscore{} should be able to use these identifiers to retrieve any particular security context.
%
In this work, as in~\cite{Norr21}, we verify that the parties agree on the values of \mCi{} and \mCr{}.

Messages 2 and 3 also serve to identify and authenticate each party to the other.
%
These messages contain long-term key identifiers (\mIdcredi{} and \mIdcredr{}), as well as authenticating information (\mAuthi{} and \mAuthr{}) which lets each party know that the other party does indeed control the long-term key provided by them.
%
The authentication information changes based on the authentication method associated with the long-term key chosen by the party in question.
%
Consider the following example scenario.
%
The initiator $I$ chooses the method $\mSigStat{}$ and sends this via $\mMethod{}$ to the responder $R$.
%
$R$ now has an option to either reject this choice of method, or to provide an identifier \mIdcredr{} for a key pair which can be used with challenge-response signatures as well as authenticating information \mAuthr{}, as dictated by the static key authentication method \mStat{}.
%
$I$ will now respond with an identifier \mIdcredi{} corresponding to a key pair which is suitable for regular signatures, and provide authentication \mAuthi{} as appropriate for the signature-based method \mSig{}.

\subsection{Key Schedule}
\label{sec:keysched}
As mentioned above, messages 2 and 3 of \mEdhoc{} contain authenticating information.
%
This contains encrypted and potentially signed MAC values, as we will describe in more detail in Section~\ref{sec:edhocauth}.
%
The keys for these are generated by using a key schedule which takes at least one DH key \mGxy{} as basic input, as shown in Figure~\ref{fig:kdfdiagram}, and is intrinsic to the functioning of \mEdhoc{}.
%
\begin{figure}[!h]
\scalebox{.785}{
\input{kdfdiagram.tex}
}
\caption{Key schedule: $P_e, P_I, P_R$ are the DH keys, \mPRKtwo, \mPRKthree, \mPRKfour are the intermediate key material, and \mKtwoe,\mKtwom, \mKthreeae, \mKthreem are the encryption keys for \mAead{} or \mXor{}. Dashed boxes are conditionals.~\cite{Norr21}}
\label{fig:kdfdiagram}
\end{figure}
%

To derive keys, \mEdhoc{} uses two functions from the \mHkdf{} interface~\cite{rfc5869}, \mHkdfExtract{} and \mHkdfExpand{}.
%
Both functions take as argument two values -- a salt, and an input. 
%
For \mHkdfExtract{}, the input is a DH key, while for \mHkdfExpand{}, it is intermediate key material.
%

As mentioned earlier, the fundamental building block for the key schedule is the ephemeral DH key \mGxy{}, which  is computed in two different ways by $I$ (as $\mDH(\mX, \mGy)$) and $R$ (as $\mDH(\mY, \mGx)$).
%
This key gives rise to intermediate keys \mPRKtwo{}, \mPRKthree{} and
\mPRKfour{}, which can be derived as part of protocol execution.
%
Each intermediate key gives rise to encryption and integrity keys (\mKtwoe, \mKtwom{}, \mKthreeae, and \mKthreem) for a particular message in the protocol.
%

In order to generate the final keys, the two \mHkdf{} algorithms use various values for salt.
%
\mPRKtwo{} is generated by the \mHkdfExtract{} algorithm while using the empty string as the salt.
%
\mPRKthree{} and \mPRKfour{} are separately generated if $R$ or $I$ uses the \mStat{} method, using the corresponding DH key as input and the previous intermediate key as salt.
%
The key $P_{R}$ is used if the responder uses the \mStat{} authentication method.
%
$P_{R}$ is the key computed as $\mDH(\mX,\ \mGr{}) = \mDH(\mPriv{R},\ \mGx)$.
%
On the other hand, the key $P_{I}$ is used if the initiator uses \mStat{}, and is computed as $\mDH(\mY,\ \mGi{}) = \mDH(\mPriv{I},\ \mGy)$.

These intermediate keys are fed into an \mHkdfExpand{} algorithm, as shown in Figure~\ref{fig:kdfdiagram}. 
%
This algorithm uses as the salt a value $\textit{th}$, which is a running hash of the information transmitted thus far as part of the protocol.
%
$\textit{th}_{i}$ denotes the hash corresponding to the $i^{\rm{th}}$ message.

At the end of a successful run of the protocol, the session key material for \mOscore{} is established as \mSessKey{}, which is a set of various keys.
%
This set always includes \mGxy{}, and if the initiator (resp. the responder) uses the \mStat{} authentication method, then it also includes $P_{I}$ (resp. $P_{R}$).
%
Once \mSessKey{} has been established, an \mHkdf{}-based key exporter named \mEdhoc-Exporter extracts the keys required for \mOscore{}.
%

\subsection{About authentication in \mEdhoc{}}
\label{sec:edhocauth}
We now describe how the authentication information is constructed, depending on which method is used.
%
For both \mStat{} and \mSig{}, the following information is used to compute \mAuthr{}: \mIdcredr{}, \mCredr{}, a transcript hash of all the communicated information thus far in the protocol, and \mADtwo{}, if included. 
%
\mAuthi{} uses the same pieces of information corresponding to the initiator role.
%
A MAC is obtained by feeding this material as additional data and the empty string as input to the Authenticated Encryption with Additional Data (AEAD) encryption algorithm as indicated in the established ciphersuite \mSuites{}. 
%
The encryption key for the AEAD algorithm is constructed, for both roles, using the ephemeral key material \mGx{}, \mGy{}, \mX{}, and \mY{}. 
%
The initiator computes $\mDH(\mX, \mGy)$ while the responder computes $\mDH(\mY, \mGx)$, and DH operations give rise to the same key under both these computations.

When $I$ uses the \mSig{} authentication method, \mAuthi{} is $I$'s signature over the MAC along with the data covered by the MAC.
%
However, when $I$ uses the \mStat{} method, \mAuthi{} is just the MAC, with one difference: the MAC key is derived using both the ephemeral key material as well as the long-term key for the initiator $\langle\mPriv{I},\ \mPub{I}\rangle$.
%
This is similar to the 1-RTT semi-static pattern in \mOptls which computes the MAC key \textsf{sfk} for the message \textsf{sfin}~\cite{DBLP:conf/eurosp/KrawczykW16}.
%
The same procedure works for $R$ as well (with the values corresponding to $R$). 
%
The abstract calculation for the authentication values is as shown in Table~\ref{tab:authvalues}~\cite{Norr21}.
%
We use $\mathit{MAC}_{I}$ (resp. $\mathit{MAC}_{R}$) to denote a MAC which uses an encryption key constructed using $\langle\mPriv{R},\ \mPub{R}\rangle$ and $\langle\mX,\ \mGx\rangle$ (resp. using $\langle\mPriv{I},\ \mPub{I}\rangle$ and $\langle\mY,\ \mGy\rangle$).
\begin{table}[ht]
\centering
\begin{tabular}{|c|c|c|}
        \hline
        \mMethod & \mAuthi & \mAuthr\\
        \hline
        \mSigSig{} & $\mSign{I}(\cdot)$ & $\mSign{R}(\cdot)$ \\
        \mSigStat{} & $\mSign{I}(\cdot)$ & $\textit{MAC}_R(\cdot)$\\
        \mStatSig{} & $\textit{MAC}_I(\cdot)$ & $\mSign{R}(\cdot)$\\
        \mStatStat{} & $\textit{MAC}_I(\cdot)$ & $\textit{MAC}_R(\cdot)$\\
        \hline
\end{tabular}
\caption{The outer functions for each method \mMethod{}~\cite{Norr21}}
\label{tab:authvalues}
\end{table}
%

In addition to this, communications 2 and 3 contain messages which are encrypted and integrity protected, as can be seen in~\ref{fig:edhocFramework}.
%
The second communication is encrypted by performing an \mXor{} operation between the plaintext and the output of the key derivation function \mHkdf{}, as in Section~\ref{sec:keysched}.
%
For the third message, encryption and integrity protection is assured by the \mAead{} algorithm which is part of the established ciphersuite \mSuites{}.

In Figure~\ref{fig:edhocsigstat} we show an example of protocol execution under the \mSigStat{} method. 
%
The figure describes in detail the various message patterns, and operations and key derivations used to construct these messages.
\vnote{Do we still have the other method diagrams that we wrote out for the original paper but did not include?}

\begin{figure}[ht]
\centering
\scalebox{.7}{
\tikzset{>=latex, every msg/.style={draw=thick}, every node/.style={fill=none,text=black}}
\begin{tikzpicture}
    \node (ini) at (0, 0) {Initiator};
    \draw [very thick] (0, -0.5) -- (0,-14.8);
    \draw [very thick] (9, -0.5) -- (9,-14.8);
    \node[below=0.5em of ini,fill=white] {$
    \begin{array}{c}
        \text{Knows}\ \langle\mPriv{I},\ \mPub{I}\rangle,\ \mIdcredi,\ \mIdcredr,\ \mADone,\ \mADthree
    \end{array}
    $};
    \node (res) at (9,0) {Responder};
    \node[below=0.5em of res,fill=white] {$
    \begin{array}{c}
        \text{Knows}\ \langle\mPriv{R},\ \mPub{R}\rangle, \ \mIdcredr,\ \mADtwo
    \end{array}$};
    \action{3.7em}{ini}{Generates $\mMethod,\ \mSuites,\ \mCi,\ \langle\mX{},\ \mGx\rangle$};
    \msg{7em}{ini}{res}{\mMsgone: \mMethod, \mSuites, \mGx, \mCi, \mADone};
    \action{8em}{res}{$
      \begin{array}{c}
          \text{Generates } \mCr,\ \langle\mY{},\ \mGy\rangle\\
          \ \ P_e = \mDH(\mY,\ \mGx{})\\
          \ \ P_R = \mDH(\mPriv{R},\ \mGx{})\\
        \mTHtwo = \mHash(\mMsgone, \langle \mCi, \mGy, \mCr \rangle)\\
        \mPRKtwo = \mHkdfExtract(\textrm{``\phantom{}''}, P_e) \\
        \mPRKthree = \mHkdfExtract(\mPRKtwo, P_R) \\
        \mKtwom = \mHkdfExpand(\mPRKthree, \mTHtwo) \\
        \mMactwo = \mAead(\mKtwom; \langle \mIdcredr, \mTHtwo, \mCredr, \mADtwo \rangle; \textrm{``\phantom{}''}) \\
        \mKtwoe = \mHkdfExpand(\mPRKtwo, \mTHtwo)
      \end{array}$};
    \msg{24em}{res}{ini}{\mMsgtwo: \mCi, \mGy, \mCr, $\overbrace{\mKtwoe\ \mXor\ \langle \mIdcredr, \mMactwo, \mADtwo \rangle}^{\mCipher}$};
    \action{25em}{ini}{$
      \begin{array}{c}
        %\mTHtwo = \mHash(\mMsgone, \langle \mCi, \mGy, \mCr \rangle) \
        \ P_e = \mDH(\mX,\ \mGy{})\\
        \mPRKtwo = \mHkdfExtract(\textrm{``\phantom{}''}, P_e) \\
        %\mKtwoe = \mHkdfExpand(\mPRKtwo,\mTHtwo)\\
        %\mGrx = \mCredr^{x} \\
        \ \ P_R = \mDH(\mX,\ \mPub{R})\\
        \mPRKfour = \mPRKthree = \mHkdfExtract(\mPRKtwo, P_R) \\
        %\mKtwom = \mHkdfExpand(\mPRKthree, \mTHtwo) \\
        \mKthreeae = \mHkdfExpand(\mPRKthree, \mTHtwo) \\
        \mTHthree = \mHash(\mTHtwo, \mCipher, \mCr)\\
        \mKthreem = \mHkdfExpand(\mPRKfour, \mTHthree) \\
        \mMacthree = \mAead(\mKthreem; \langle \mIdcredi, \mTHthree, \mCredi, \mADthree \rangle; \textrm{``\phantom{}''}) \\
        \mSigthree = \mSign{I}(\langle \langle \mIdcredi, \mTHthree, \mCredi, \mADthree \rangle, \mMacthree \rangle)
      \end{array}$};
    \msg{38.5em}{ini}{res}{$\mMsgthree: \mCr, \mAead(\mKthreeae; \mTHthree; \langle \mIdcredi, \mSigthree, \mADthree \rangle$)};
    \action{40em}{res}{$
    \begin{array}{c}
        \mTHthree = \mHash(\mTHtwo, \mCipher, \mCr)\\
        \mKthreem = \mHkdfExpand(\mPRKthree, \mTHthree) \\
        \mKthreeae = \mHkdfExpand(\mPRKthree, \mTHthree)
    \end{array}$};
    \draw [line width=1mm] (-2,-14.8) -- (2,-14.8);
    \draw [line width=1mm] (7,-14.8) -- (11,-14.8);
    \end{tikzpicture}
}
    \caption{The \mSigStat{} method for \mEdhoc{}. $\langle\cdot\rangle$ denotes a tuple, and the hash function \mHash{} is as established in the ciphersuite \mSuites{}.~\cite{Norr21}}
\label{fig:edhocsigstat}
\end{figure}
%------------------------------------------------------------------------- sub