### 2022-07-21 paper.tex
[add] updates and fixes to section 6

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}

\usepackage{epsfig}
\usepackage{subcaption}
\usepackage{calc}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{multicol}
\usepackage{pslatex}
\usepackage{apalike}

\usepackage[scaled=0.8]{helvet}    % Less huge \textsf{functionName}
\usepackage[misc,geometry]{ifsym} % for letter symbol
\usepackage{graphicx}
\usepackage[dvipsnames]{xcolor}
\usepackage{tikz}
\usetikzlibrary{trees,snakes,arrows}
\usetikzlibrary{shapes,chains}
\usetikzlibrary{positioning}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue
}
\usepackage[nospace]{cite}

\input{macros}

\title{Formal Analysis of EDHOC Key Establishment for Constrained IoT Devices}
\author{
        Karl Norrman and 
        Vaishnavi Sundararajan and
        Alessandro Bruni
}

\begin{document}
\maketitle
%
\begin{abstract}
Given how common IoT devices that use constrained resources are becoming today, the need of the hour is communication protocols which can operate securely under such limitations.
%
For a few years, the Internet Engineering Task Force (IETF) has been working to standardize \mEdhoc{}, an authenticated key establishment protocol for such constrained IoT devices.
%
The first version of \mEdhoc{} was proposed in 2016.
%
In 2018, Bruni et al.~\cite{DBLP:conf/secsr/BruniJPS18} used the \mProverif{} tool~\cite{DBLP:conf/csfw/Blanchet01} to formally analyze an early version of \mEdhoc{}, which had only two key establishment methods.  
%
%By now, the protocol has been fleshed out much more, and has been augmented with multiple new key establishment methods.
By 2021, the protocol had been fleshed out much more, with multiple new key establishment methods, and this version was formally analyzed using the \mTamarin{} prover~\cite{DBLP:conf/cav/MeierSCB13} in~\cite{Norr21}.
%
Here, we build on this work, and use \mTamarin{} to formally analyze the key establishment methods in the current version of \mEdhoc{}, as well as discuss some ramifications of the choices made while designing the protocol.
\vnote{We need to see what more we can say here, about the new stuff we're potentially adding}
\end{abstract}
%
%-------------------------------------------------------------------------- sec
\section{\uppercase{Introduction}}
\label{sec:introduction}
\input{intro}

%-------------------------------------------------------------------------- sec
\section{\uppercase{The \mEdhoc{} Protocol}}
\label{sec:edhoc}
\input{protocol}

%-------------------------------------------------------------------------- sec
\section{\uppercase{Formalization and Results}}
\label{sec:formalization}
\input{formalization}

%-------------------------------------------------------------------------- sec
\section{\uppercase{Error handling}}
\label{sec:errorHandling}
Section 6 of the current \mEdhoc{} specification~\cite{draft-ietf-lake-edhoc-15} states that error messages can be sent at any time and by any party.
%
There are three types of error messages: type 0 is used to represent success, type 1 represents a generic error message used for debugging purposes, and type 2 represents the failure to negotiate a common ciphersuite. 
%
Type 2 messages carry a list of supported ciphersuites by the responder.
%
The contents and semantics of type 1 messages are dependent on the implementation.
%
Consequently, type 1 error messages provide a generic message passing mechanism, albeit without a predefined semantics.
%
We now discuss some issues with this specification of error messages.

\subsection{Lack of connection identifiers}
Error messages do not include the optional connection identifiers \mCi{} or \mCr{}.
%
This is potentially problematic for connection-less transport layers, where these identifiers may be used to correlate messages.

\subsection{Improper handling}
\mEdhoc{} provides an error message mechanism, but little to no guidance on how it should be used safely.
%
For example, assume that an application using \mEdhoc{} logs error messages on a finite log with rotation.
%
If the log is used for anomaly detection or detection of sensitive events, then an adversary can simply inject error messages and fill the log until the sensitive event is overwritten.
%
In such situations, implementors should do proper log-separation and log management.

\subsection{Unspecified semantics for debugging messages}
Section 6.2 gives an example of sending a type 1 error message with the information string "Method not supported".
%
Note that \mEdhoc{} does not include a mechanism to support the negotiating of the connection method.
%
The initiator selects the method as part of the first message, and if the responder replies with a type 1 error message along with the information string "Method not supported", the standard does not tell an implementation how to act upon this information string for negotiation purposes.
%
An adversary can fake an error message, and there is no algorithm that is specified to ensure that a common available authentication mechanism is always selected.
%
In particular, note that Section 5.1 in the specification states that if processing fails for some reason, then "typically" an error message is sent, processing is terminated and protocol state is discarded. 
%
But negotiation needs us to remember state.

With respect to ciphersuite negotiation, Section 5.2.2 of the specification also states ``The Initiator MUST NOT change the supported cipher suites and
the order of preference in SUITES\_I based on previous error messages.'' 
%
However, \mSuites is sent in the clear, and therefore can be modified by the adversary.
%
It is important to consider what happens if the order of preference in \mSuites is changed either by the adversary, or by the initiator themselves for some other reason.
%
It might be helpful to include a flag in \mMsgone to indicate to the responder
that a (modified) ciphersuite ordering depends on that from a previous
execution.

%-------------------------------------------------------------------------- sec
\section{\uppercase{Improvements to the specification}}
\label{sec:discussion}
The \mEdhoc{} \mSpec{} has many points where it can be improved.
%
We discovered a few of these issues in~\cite{Norr21}, but in this work,
we also discuss a couple of new issues.
%
We communicated the points discussed in~\cite{Norr21} to the authors of the
\mSpec{}.
%
Here we also list the protocol developers' response to being told of these
ambiguities in the \mSpec{}.

%-------------------------------------------------------------------------- sub
\subsection{Unclear Intended Use}
\label{sec:unclearProtocolUse}
%
There are several security goals listed in the \mSpec{}.
%
However, these goals are informal and imprecisely specified, and the lack of
intended usage makes them difficult to interpret.
%
Since we have little indication about use case restrictions, it is also hard
to evaluate whether the security goals are the most relevant ones.
%

In order to identify security goals that might be relevant
we made up some user stories corresponding to typical use cases,
which helped us identify some subtle points of concern in \mEdhoc.
%
We communicated these points to the authors of \mEdhoc{}.
%

%----------------------------------------------------------------------- subsub
\subsubsection{Non-Repudiation.}
Access control mechanisms can vary depending on what the underlying application
is.
%
A nuclear power-plant might need to keep track of who enters and leaves, but a
coffee machine that logs every user and their coffee preference might lead to a
privacy concern.
%
This simple thought experiment allowed us to identify that the \mEdhoc{}
\mSpec{} did not even consider non-repudiation.
%
When we pointed this out to the authors of the \mSpec, they recognized this
concern, and added a discussion about how the various \mEdhoc{} methods
satisfy (or not) non-repudiation.
%

%----------------------------------------------------------------------- subsub
\subsubsection{Unintended Peer Authentication.}
\label{sec:unintendedPeerAuth}
In Section~3.2 in the \mSpec{}, the authors state that parties executing
\mEdhoc{} must be configured such that they only run \mEdhoc{} with a
restricted (according to some reasonable policy) set of peers.
%
However, the responsibilities of verifying which identity was authenticated is
not clearly split between the application and \mEdhoc{} and therefore an attack
similar in effect to UKS is possible.
%

Suppose a person configures the restriction policy such that every device in
their home is allowed to set up an \mEdhoc{} session with any other device in
the home.
%
However, the adversary has managed to gain control of one of these devices (say
$A$).
%
Device $A$ is part of the allowed group of devices, so $B$ will accept an
\mEdhoc{} run with $A$, even though $A$ is now compromised.
%
If the application in $B$ tasks the \mEdhoc{} implementation to establish a
security context with $C$, $B$ will send an initial message to $C$.
%
However, if $A$ responds to the initial message before $C$ can, the \mSpec{}
does not require $B$ to verify that \mIdcredr{} matches what is expected for
$C$.
%
A valid \mEdhoc{} implementation may then look up the appropriate credentials
for $A$, complete the execution and deliver the security context to the
application.
%
The application gets no indication of that the security context is shared with
$A$ rather that $C$ as intended.
%

A straight forward check that the received \mIdcredr{} matches the expected
identity thwarts the problem.
%
We included this check in our model.
%
We communicated this to the authors of \mEdhoc{}, who in the latest version of
the \mSpec{}, added that \mEdhoc{} makes \mIdcredr{} available to the
application, and added a new appendix D describing credential validations that
the application must take care of.
%

%----------------------------------------------------------------------- subsub
\subsubsection{Denial of Service.}
Denial of Service (DoS) attacks for IoT protocols are well-studied,
but DoS aspects particular to \mEdhoc{} is not.
%
The \mSpec{} does not include any specific measures for countering such attacks,
but assumes the application takes care of it.

Section 8.7 of the latest version of the \mEdhoc{}
\mSpec{}~\footnote{draft-ietf-lake-edhoc-15}
gives two recommendations for countering DoS attacks.
%
First, \mEdhoc{} relies on lower layers to mitigate DoS, and an example of DoS
countermeasures is given: checking the return address upon receipt of the first
message.
%
Second, a recommendation is given that applications may try to determine (in
some way) that a seemingly valid message is in fact probably a forgery and
should be ignored or cause the protocol to abort.
%

No recommendation is given on how to minimize DoS effects stemming from the use
of error messages, which can be used in a relatively arbitrary way in \mEdhoc{}
(see Section~\ref{sec:errorHandling}).
%
We now elaborate on some considerations that the application developers need to
take into account.
%

As in the IKEv2 DDoS analysis by~\cite{rfc8019}, the responder in \mEdhoc{}
is more vulnerable than the initiator.
%
The key similarity is that the initiator enjoys identity protection in the
first message.
%
None of the elements in the first \mEdhoc{} message can be proved by the
responder to be invalid, as long as the format is correct and the chosen
elliptic curve points lie on the chosen curve.
%
Additionally, execution may require the responder to obtain and verify
certificate chains.
%
For example, the content of the \mADone{} information element may need the
responder to take further action.
%
Since \mEdhoc{} is designed for constrained devices, this asymmetry needs to be
considered even more carefully than for IKEv2~\cite{rfc8019}.
%

The asymmetry between the adversary's effort towards the first message and that
of the responder may hence be significant and the possibility for applications
to select which entity acts as initiator and which acts as responder may
important.
%
\mEdhoc{} is profiled to work in conjunction with \mCoap{} and
\mOscore{}~\cite{ietf-core-oscore-edhoc-03}.
%
The profiling locks the \mEdhoc{} initiator role to the \mCoap{} client role and
hence reduces the possibilities to change the \mEdhoc{} roles according to DoS
vulnerability preferences.
%

Reflection attacks are another risk.
%
These are attacks where an adversary spoofs the source address of a victim in a
message and thereby tricks the responder into sending a message to the victim.
%
\mEdhoc{} suggests using a return path reachability method similar to IKEv2.
%
Note that even such reachability mechanisms can easily be used to make a
responder send reachability requests to arbitrary addressable targets.
%
However, such mechanisms might require that unsolicited reachability requests be
discarded, limiting the effects of anonymous DoS attacks.
%
In situations where the responder does not need to perform additional
communications to verify certificates etc, a reachability mechanism may be more
expensive in terms of time and storage compared to continuing with the protocol
while keeping a half-open state for a period of time.
%
However, if round-trip times are large, and half-open states hence are kept by
the receiver for a long time, a reachability mechanism can trade communication
for storage.
%

\mEdhoc{} responders querying external servers for certificate revocation checks
or lookups may casue systems-level issues if many distributed \mEdhoc{}
responders frequently queries such servers when DoSed by clients.
%

An adversary may also send forged error messages (see
Section~\ref{sec:errorHandling}), and specifically reject proposed ciphersuites.
%
The \mEdhoc{} \mSpec{} recommends the initiator to not try that ciphersuite with
the responder again in a new session.
%
If the adversary rejects all options for an initiator this way, it can prevent
communication between the initiator and that responder until the initiator's
cache is cleared.
%
Potentially, an adversary could lock a party of a group out of communication
altogether this way.
%

%------------------------------------------------------------------------- sub
\subsection{Unclear Security Model}
In addition to unclear contexts in which
\mEdhoc{} can be used, the \mSpec{} also does not provide enough detail about
adversary capabilities.
%
The four different cryptographic cores included in \mEdhoc{} are based on
academic protocols which are designed to work with highly specific but
potentially different adversary models.
%
%For example, \mOptls{}, which essentially has the same core as \mStat{}, is
%shown to be secure in the CK model~\cite{DBLP:conf/crypto/CanettiK02}, which
%explicitly separates long-term key storage from that of the ephemeral keys and
%session state.
%%
%This fits in well with our use of TEEs.
Since \mSigma{} cannot protect against compromised ephemeral keys, the authors
of \mEdhoc{} felt that considering the compromise of ephemeral keys separately
from that of long-term keys was not required~\cite{personalCommunication}.
%
The reason was presumably based on the fact that the \mSigSig{} method was
modeled closely on the \mSigmaI{} variant of \mSigma{}, and that it would be
preferable to obtain a homogeneous security level among the \mEdhoc{}
methods.~\cite{Norr21}
%
However, this only holds true for situations where one is only interested in the
session key confidentiality for an ongoing session.
%
Secure modules, in addition to being able to store long-term keys, are useful in
other ways too.
%
They also provide weak Post-Compromise Security (PCS) guarantees as discussed
above.
%
Upon discussion with the authors, they included recommendations on how to store
long-term keys, and how to perform operations on these keys inside a secure
module.
%

%-------------------------------------------------------------------------- sub
\subsection{Session Key Material}
\label{sec:sessionKeyMaterial}
As we have seen, \mEdhoc{} establishes some
session key material, which can be fed into the \mEdhoc{}-Exporter to derive
session keys.
%
This key material is directly influenced by \mGxy{} and a party's secret static
long-term key, if they use \mStat{}.
%
As mentioned earlier, mutual explicit injective agreement cannot be obtained for
$P_{I}$.
%
This might prove to be a problem, and we proposed three potential alternatives
for addressing this to the \mEdhoc{} authors.
%
\begin{itemize}
    \item Include \mIdcredi{} or a hash thereof in the first and
    second messages:
%
    This increases message size and also leaks the initiator's identity.
%
    \item Derive session key material without using $P_{I}$:
%
    This deviates from the structure of protocols like \mOptls{} from which the
    \mStat{}-based methods draw heavy inspiration.
%
    In those protocols, $P_{I}$'s presence in the computation of session key
    material is crucial to obtaining resistance against the compromise of the
    initiator's ephemeral key.
%
    \item Include a fourth message from the
    responder to the initiator which contains a MAC based on information
    obtained using $P_{I}$:
\end{itemize}
%
While the last option increases the number of messages in the protocol, the
\mEdhoc{} authors decided to go with this option for situations where it is
necessary to provide explicit injective agreement on $P_{I}$.
%

%-------------------------------------------------------------------------- sec
\section{\uppercase{Conclusions and Future Work}}
\label{sec:conclusions}
\label{sec:newdrafts}
We formally modeled all four
methods of the \mEdhoc{} \mSpec{} using \mTamarin.
%
We formulated several important security properties and identified precise
adversary models in which we verified these.
%
The properties are shown in Table~\ref{tab:props}.
%
Mutual injective agreement covers the parameters $S_P$:
responder identity, roles, session key material (except for $P_I$ when
initiator uses the \mStat{} authentication
method), context identifiers \mCi{} and \mCr, and cipher suites \mSuites.
%
The responder in addition is ensured agreement on the initiators identity and
$P_I$, i.e., on the set $S_F$.
%
Implicit agreement covers all previously mentioned parameters for both peers.
%
All measurements were taken on a server with 2 Intel(R) Xeon(R) Gold
6242 CPU @ 2.80GHz with a total of 32 cores / 64 threads and reserved
256 GiB of memory for the scheduled job (the machine has a total of
384 GiB installed).
%
Time measurements are rounded to the nearest second and we report both
real and user times (i.e. time to complete the task and total CPU time
spent for the task).
%
Mutual entity authentication, UKS- and KCI resistance can be inferred
from the verified properties.
\knote{In Section 3.3 we say that injective agreement does \emph{not} hold for
    all methods. It is a little contradictory to the table below where we
    exclude $P_I$. It would be better to show it as not holding when $P_i$ is
    included (both for consistency and for clarity).
}
%
\begin{table*}[t]
        \centering
          \label{tab:props}
          \resizebox{\textwidth}{!}{%
            \begin{tabular}{l|c|c|c|c}
              \textbf{Lemma $\backslash$ Method} & \textbf{SIG-SIG} & \textbf{SIG-STAT} & \textbf{STAT-SIG} & \textbf{STAT-STAT}\\
              \hline
              \textbf{Injective agreement / I} & $S_F$ & $S_F$ & $S_P\ddagger$ & $S_P\ddagger$\\
              \textbf{Injective agreement / R} & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
              \textbf{Time (real/user)} & 0m27s / 11m17s & 9m27s / 239m49s & 1m21s / 35m36s & 18m16s / 499m2s\\
              \textbf{Implicit agreement / I} & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
              \textbf{Implicit agreement / R} & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
              \textbf{Time (real/user)} & 2m37s / 65m40s & 5m28s / 131m12s $\dagger$ & 8m54s / 212m21s & 7m37s / 188m46s $\dagger$\\
              \textbf{Session key secrecy} & \cm & \cm & \cm & \cm\\
              \textbf{Time (real/user)} & 0m41s / 16m4s & 1m52s / 45m27s & 6m31s / 156m26s & 9m47s / 232m2s\\
              \textbf{Secrecy of AD3} & \cm & \cm & \cm & \cm\\
              \textbf{Integrity of AD3} & \cm & \cm & \cm & \cm\\
              \textbf{Time (real/user)} & 1m40s / 38m57s & 5m53s / 138m43s & 3m35s / 85m12s & 45m6s / 1029m1s\\
            \end{tabular}}\\[1em]

          \caption{Verified properties. $S_P$ contains roles,
            responder identity, session key material (excluding
            $P_I$), \mCi, \mCr, and \mSuites. $S_F$ is $S_{P}$,
            the initiator identity, and $P_I$.\\
            $\dagger$: For SIG-STAT and STAT-STAT implicit
            authentication using the full XOR model the verification
            was killed at 14246m43s and 2057m47s of computation time,
            respectively. The time reported is for verification under
            the simplified XOR modeling of message 2.\\
            $\ddagger$: When the initiator is using STAT mode we
            cannot get explicit agreement on the intiator's own key
            material, therefore in these two cases we check injective
            agreement for all key material except $P_I$.\\
          }
        % \begin{tabular}{|l|c|c|c|c|}
        %         \hline
        %         & \mSigSig & \mSigStat & \mStatSig & \mStatStat \\
        %         \hline
        %         Injective agreement for I & $S_F$ & $S_F$ & $S_P$ & $S_P$\\
        %         Injective agreement for R & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
        %         Implicit agreement for I & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
        %         Implicit agreement for R & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
        %         PFS for session key material & \cm & \cm & \cm & \cm\\
        %         \hline
        % \end{tabular}
\end{table*}

Further, we identified a situation where initiators may establish an \mOscore{}
security context with a different party than the application using \mEdhoc{}
intended, and proposed a simple mitigation.
%
We discussed how the IETF may extract and better define security properties to
enable easier verification.

We verified each method in isolation, and leave as future work to verify whether
the methods are secure under composition.

%\subsection{Newer Versions of the Specification}
%\label{sec:newdrafts}
In this work, we have analyzed the \mEdhoc{} version as of July
2020~\cite{our-analysis-selander-lake-edhoc-00}.
%
There are newer versions and our analysis applies to at least the version from
February 2021~\cite{latest-ietf-lake-edhoc-05}.
%
The latest version is \emph{draft-ietf-lake-edhoc-15}
%
\knote{we are already citing too many different versions, so I don't think we
    should add this to the bib. Potentially, we should tomorrow instead remove
    and just mention some of the later versions.}
%

%-------------------------------------------------------------------------- ack
\section*{ACKNOWLEDGEMENTS}
This work was partially supported by
the Wallenberg AI, Autonomous Systems and Software Program (WASP) funded by
the Knut and Alice Wallenberg Foundation.
%
We are grateful to G\"oran Selander, John Mattsson and Francesca Palombini for
clarifications regarding the specification.
%

%-------------------------------------------------------------------------- bib
\bibliographystyle{apalike}
{\small
    \bibliography{refComp}
}
\end{document}